---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Build Status](https://travis-ci.org/graggsd/tidytableR.svg?branch=master)](https://travis-ci.org/graggsd/tidytableR)
[![Build status](https://ci.appveyor.com/api/projects/status/nsalen7ay5tuwpqd/branch/master?svg=true)](https://ci.appveyor.com/project/graggsd/tidytabler/branch/master)

# tidytableR

The goal of tidytableR is to is to reshape the output of an analysis
from a tidy format into a for suitable for presentation. It allows users to 
map column names from an input data frame to elements of an output table.

The package was developed from code meant to act as a wrapper script for the 
`htmlTable` function in the 
[htmlTable](https://cran.r-project.org/web/packages/htmlTable/index.html)
package. Therfore, many of the variable names and symantics are related to
those used in the `htmlTable` package.

## Installation

You can install tidytableR from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("graggsd/tidytableR")
```

