---
title: "Basic use of tidytableR"
author: "Stephen Gragg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic use of tidytableR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

`tidy_htmlTable` acts as a wrapper function for the `htmlTable`
function allowing columns to be mapped from the input data to specific htmlTable 
parameters in a manner similar to ggplot2.

`tidy_df` acts in a similar manner, except that it outputs a `data.frame` to 
make saving its output into tabular files easier.

# Example Use

## Prepare Data

We'll begin by turning the `mtcars` data into a tidy dataset. The 
`gather` function is called to collect 3 performance metrics into a pair 
of key and value columns.

```{r, message=FALSE, warning=FALSE}
library(tidytableR)
library(magrittr)
library(dplyr)
library(tidyr)
library(tibble)
```

```{r}
td <- mtcars %>%
    rownames_to_column %>%
    select(rowname, cyl, gear, hp, mpg, qsec) %>%
    gather(per_metric, value, hp, mpg, qsec)
```

Now we will compute 4 summary statistics for each of the 3 performance metrics.
This will be further grouped by number of cylinders and gears.

```{r}
tidy_summary <- td %>%
    group_by(cyl, gear, per_metric) %>% 
    summarise(Mean = round(mean(value), 1),
              SD = round(sd(value), 1),
              Min = round(min(value), 1),
              Max = round(max(value), 1)) %>%
    gather(summary_stat, value, Mean, SD, Min, Max) %>% 
    ungroup %>% 
    mutate(gear = paste(gear, "Gears"),
           cyl = paste(cyl, "Cylinders"))
```

Though in this format, `tidy_summary` is easy to pass onto functions within
`ggplot2`, with 86 rows, it is not conducive to presentating as a table. Though
`tidyr` may be used to create a wide format, this will take some work, and there 
will still be additional steps if one wishes to present the data using the
`htmlTable` package. The `tidytableR` package was created to shortcut these
steps.

## `tidy_htmlTable`

Columns from `tidy_summary` may be mapped directly to arguments within the 
`htmlTable` function similar to the way that columns are mapped to elements
of a plot within `ggplot2`. That said, to use functions within this package 
appropriately, it is important to be familiar with the `htmlTable` package.

Note that, at minimum, the `value`, `rnames`, and `header` arguments must be
specified.

Also, the combination of `rnames` and `header` 
(additionally, `rgroup`,  `tspanner`, `cgroup1`, and `cgroup2`, if they 
are specified by the user) must form a unique key for each row. 
This will most often occur naturally if the the input data has been created 
`dplyr::summarise` or `tidyr::gather`, which are the primary use cases for these
functions.

### Example 1

```{r}
tidy_summary  %>% 
    tidy_htmlTable(header = "gear",
                   cgroup1 = "cyl",
                   cell_value = "value", 
                   rnames = "summary_stat",
                   rgroup = "per_metric")
```

### Example 2

```{r}
tidy_summary  %>% 
    tidy_htmlTable(header = "summary_stat",
                 cgroup1 = "per_metric",
                 cell_value = "value", 
                 rnames = "gear",
                 rgroup = "cyl")
```

## `tidy_df`

Often, a tabular data file will be desired for sharing or manual editing.
`tidy_df` has the same functionality as `tidy_htmlTable`, but returns 
a `data.frame` instead of html code.

### Without collapsing

By default, the output of `tidy_df` will include extra columns on the left 
margin of the output `data.frame` to designate values for `tspanner`, 
`rgroup`, and `rnames` (ordered from outside to inside respectively), and the
top margin to designate `cgroup2`, `cgroup1`, and `header` values
(ordered from outside to inside respectively).

```{r}
tidy_summary  %>%
    # So that output may be displayed appropriately
    filter(cyl != "6 Cylinders") %>% 
    tidy_df(header = "gear",
            cgroup1 = "cyl",
            cell_value = "value", 
            rnames = "summary_stat",
            rgroup = "per_metric")
```

### With collapsing

As an alternative, a character string may be specified in the `collapse` 
argument, which will result in `tspanner`, `rgroup`, and `rnames` being 
combined into a single term `collapse` and included as a single column on the 
far left margin of the table, and `cgroup2`, `cgroup1`, and `header` receiving 
the same treatment, but being used to name columns of the output `data.frame`.

```{r}
tidy_summary  %>%
    # So that output may be displayed appropriately
    filter(cyl != "6 Cylinders",
           cyl != "8 Cylinders", 
           gear != "5 Gears") %>% 
    tidy_df(header = "gear",
            cgroup1 = "cyl",
            cell_value = "value", 
            rnames = "summary_stat",
            rgroup = "per_metric",
            collapse = ".")
```

